---
title: "explore titanic"
author: "Roland Krasser"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{explore_titanic}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## How to explore the Titanic dataset using the explore package.

The explore package simplifies Exploratory Data Analysis (EDA). Get faster insights with less code!

The titanic dataset is available in base R. The data has 5 variables and only 32 rows. Each row does NOT represent an observation. It is not tidy, instead the data set contains Frequencies! And it is not a data frame, therefore we we need to convert it first.

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(explore)
titanic <- as.data.frame(Titanic)
```

### Explore dataset

```{r message=FALSE, warning=FALSE}
titanic %>% describe()
```

All variables are categorical except Freq, representing the frequency.

The data look like this:

```{r message=FALSE, warning=FALSE}
titanic %>% head(10)
```

### Explore variables

As the normal explore() function of the {explore} package expects a tidy dataset (each row is an observation), we need add the parameter n (number of observations)

```{r message=FALSE, warning=FALSE, fig.height= 2.5, fig.width=4}
titanic %>% explore(Class, n = Freq)
```

We get the exact numbers by using count() together with the wt-parameter (weight)

```{r message=FALSE, warning=FALSE}
titanic %>% count(Class, wt = Freq)
```

```{r message=FALSE, warning=FALSE, fig.height= 2.5, fig.width=4}
titanic %>% explore(Sex, n = Freq)
```

```{r message=FALSE, warning=FALSE, fig.height= 2.5, fig.width=4}
titanic %>% explore(Age, n = Freq)
```

```{r message=FALSE, warning=FALSE, fig.height= 2.5, fig.width=4}
titanic %>% explore(Survived, n = Freq)
```

### Survived?

Now we want to check the relation between variables and Survived. We can use the explore() function with Survived as target.

```{r message=FALSE, warning=FALSE, fig.height= 2.5, fig.width=4}
titanic %>% explore(Class, target = Survived, n = Freq, split = FALSE)
```

To get a better feeling of the relationship between Class and Survived, we switch to percentage and split the target into sperate bars. We can do that by using split = TRUE (which is default).

```{r message=FALSE, warning=FALSE, fig.height= 2.5, fig.width=4}
titanic %>% explore(Class, target = Survived, n = Freq, split = TRUE)
```

Now we get a plot, where each color sum to 100%. So a big difference in bar length indicates an important relationship between the two variables. In this case, passengers of 1st Class had the highest probability to survive.

```{r message=FALSE, warning=FALSE, fig.height= 2.5, fig.width=4}
titanic %>% explore(Sex, target = Survived, n = Freq)
```

Female are much more likely to survive!

```{r message=FALSE, warning=FALSE, fig.height= 2.5, fig.width=4}
titanic %>% explore(Age, target = Survived, n = Freq)
```

Child had an advantage to survive.

Now we can create a simple decision tree. As we have count-data we need to pass parameter n.

```{r}
titanic %>% explain_tree(target = Survived, n = Freq)
```

We see that Sex and Class can give a good explanation who are more likely to survive.

* Sex = Male: 21% survived
* Sex = Female & Class = 3rd: 46% survived
* Sex = Female & Class <> 3rd: 93% survived

### Other correlations

```{r message=FALSE, warning=FALSE, fig.height= 2.5, fig.width=4}
titanic %>% explore(Age, target = Class, n = Freq)
```

Child are unlikely in the 1st class! And all Crew members are adult as expected.

```{r message=FALSE, warning=FALSE, fig.height= 2.5, fig.width=4}
titanic %>% explore(Sex, target = Class, n = Freq)
```

Almost no female Crew members! Female tend to have better Class!
