---
title: "Explain"
author: "Roland Krasser"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Explain}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The explore package offers a simplified way to use machine learning to understand and explain patterns in the data.

* `explain_tree()` creates a decision tree. The target can be binary, categorical or numerical
* `explain_forrest()` creates a random forest. The target can be binary, categorical or numerical
* `explain_logreg()` creates a logistic regression. The target must be binary

We use synthetic data in this example

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(explore)

data <- create_data_buy(obs = 100)
glimpse(data)
```

### Decision Tree

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
data %>% explain_tree(target = buy)
```

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
data %>% explain_tree(target = mobiledata_prd)
```

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
data %>% explain_tree(target = age)
```

#### Random Forrest

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
data %>% explain_forest(target = buy, ntree = 100)
```

#### Logistic Regression

```{r message=FALSE, warning=FALSE}
data %>% explain_logreg(target = buy)
```
