---
title: "explore_penguins"
author: "Roland Krasser"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{explore_penguins}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## How to explore the penguins dataset using the explore package.

The explore package simplifies Exploratory Data Analysis (EDA). Get faster insights with less code! We will use < 10 lines of code and just 6 function names to explore penguins:

| function       | package   | description                             |
|----------------|-----------|-----------------------------------------|
| library        | {base}    | load a package                          |
| filter()       | {dplyr}   | subset rows using column values         |
| describe()     | {explore} | describe variables of the table         |
| explore()      | {explore} | explore graphically a variable          |
| explore_all()  | {explore} | explore all variables of the table      |
| explain_tree() | {explore} | explain a target using a decision tree  |

The penguins dataset comes with the palmerpenguins package. It has 344 observations and 8 variables. (https://github.com/allisonhorst/palmerpenguins)

So we have to load the palmerpenguins package. Furthermore, we use the packages {dplyr} for filter() and |> and {explore} for data exploration.

```{r message=FALSE, warning=FALSE}
library(palmerpenguins)
library(dplyr)
library(explore)
```

### Describe variables

```{r message=FALSE, warning=FALSE}
penguins |> describe()
```
There are some NA-values (unknown values) in the data. The variable containing the most NAs is sex. flipper_length_mm and others contain only 2 observations with NAs.

### Data cleaning

We use only penguins with known flipper length for the data exploration!

```{r}
data <- penguins |> 
  filter(flipper_length_mm > 0)
```

We reduced the penguins from 344 to 342.

### Explore variables

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=total_fig_height(data, size = 2.5)}
data |> 
  explore_all()
```

### Which species?

What is the relationship between all the variables and species?

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=total_fig_height(data, var_name_target = "species", size = 2.2)}
data |> 
  explore_all(target = species)
```

We already see some strong patterns in the data. flipper_length_mm seperates species Gentoo, bill_length_mm seperates species Adelie from Chinstrap. And we see that Chinstrap and Gentoo are located on seperate islands.

Now we explain species using a decision tree:

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
data |> explain_tree(target = species)
```

We found an easy explanation how to find out the species by just using flipper_length_mm and bill_length_mm.

* If flipper_legnth_mm >= 207, it is a Gentoo penguin (95% right)
* If flipper_length_mm < 207 and bill_length_mm < 43, it is a Adelie penguin (97% right)
* If flipper_length_mm < 207 and bill_length_mm >= 43, it is a Chinstrap penguin (92% right)

Now let's take a closer look to these variables:

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
data |> explore(flipper_length_mm, bill_length_mm, target = species)
```

The plot shows a not perfect but good seperation between the 3 species!
