---
title: "explore"
author: "Roland Krasser"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{explore}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Interactive data exploration with one line of code or use a easy to remember set of tidy functions for exploratory data analysis. Introduces two main verbs. **describe()** to describe a variable or table, **explore()** to grafically explore a variable or table.

explore package on Github: [https://github.com/rolkra/explore](https://github.com/rolkra/explore)

As the explore-functions fits well into the tidyverse, we load the dplyr-package as well.

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(explore)
```

### Interactive data exploration

Explore your dataset (in this case the iris dataset) in one line of code:

```{R eval=FALSE, echo=TRUE}
explore(iris)
```

A shiny app is launched, you can inspect individual variable, explore their relation to a binary target, grow a decision tree or create a fully automated report of all variables with a few "mouseclicks".
 
You can choose each variable containng 0/1, FALSE/TRUE or "no"/"yes" as a target. As the iris dataset doesn't contain a binary target, we create one:

```{R eval=FALSE, echo=TRUE}
iris$is_versicolor <- ifelse(iris$Species == "versicolor", 1, 0)
iris %>% explore()
```

### Report variables

Create a rich HTML report of all variables with one line of code:

```{R eval=FALSE, echo=TRUE}
# report of all variables
iris %>% report(output_file = "report.html", output_dir = tempdir())
```

Or you can simply add a target and create the report:

```{R eval=FALSE, echo=TRUE}
# report of all variables and their relationship with a binary target
iris$is_versicolor <- ifelse(iris$Species == "versicolor", 1, 0)
iris %>% 
  report(output_file = "report.html", 
         output_dir = tempdir(),
         target = is_versicolor)

```

### Grow a decision tree

Grow a decision tree with one line of code:

```{r fig.height=3, fig.width=6, message=FALSE, warning=FALSE, include=FALSE}
iris$is_versicolor <- ifelse(iris$Species == "versicolor", 1, 0)
```
```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
iris %>% select(-Species) %>% explain_tree(target = is_versicolor)
```

You can control the growth of the tree using the parameters `maxdepth`, `minsplit` and `cp`.

### Explore variables

Explore a variable with one line of code. You don't have to care if a variable is numerical or categorical.

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
iris %>% explore(Species)
```

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
iris %>% explore(Sepal.Length)
```

### Explore variables with a target

Explore a variable and its relationship with a binary target with one line of code. You don't have to care if a variable is numerical or categorical.

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
iris %>% explore(Species, target = is_versicolor)
```

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
iris %>% explore(Sepal.Length, target = is_versicolor)
```

### Explore multiple variables

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
iris %>% 
  select(Sepal.Length, Sepal.Width) %>% 
  explore_all()
```

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
iris %>% 
  select(Sepal.Length, Sepal.Width, is_versicolor) %>% 
  explore_all(target = is_versicolor)
```

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
iris %>% 
  select(Sepal.Length, Sepal.Width, is_versicolor) %>% 
  explore_all(target = is_versicolor, density = FALSE)
```

### Explore correlation between two variables

Explore correlation between two variables with one line of code:

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
iris %>% explore(Sepal.Length, Petal.Length)
```

You can add a target too:

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
iris %>% explore(Sepal.Length, Petal.Length, target = is_versicolor)
```

### Other options

If you use explore to explore a variable and want to set lower and upper limits for values, you can use the `min_val` and `max_val` parameters. All values below min_val will be set to min_val. All values above max_val will be set to max_val.

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
iris %>% explore(Sepal.Length, min_val = 4.5, max_val = 7)
```


`explore` uses auto-scale by default. To deactivate it use the parameter `auto_scale = FALSE`

You can force the explore-function to use bars with the parameter `density = FALSE`

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
iris %>% explore(Sepal.Length, density = FALSE, auto_scale = FALSE)
```

### Describing data

Describe your data in one line of code:

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
iris %>% describe()
```

The result is a data-frame, where each row is a variable of your data. You can use `filter` from dplyr for quick checks:

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
# show all variables that contain less than 5 unique values
iris %>% describe() %>% filter(unique < 5)
```

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
# show all variables contain NA values
iris %>% describe() %>% filter(na > 0)
```

You can use `describe` for describing variables too. You don't need to care if a variale is numerical or categorical. The output is a text.

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
# describe a numerical variable
iris %>% describe(Species)
```

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
# describe a categorical variable
iris %>% describe(Sepal.Length)
```

### Data Dictionary

Create a Data Dictionary of a dataset (Markdown File data_dict.md)

```{R eval=FALSE, echo=TRUE}
iris %>% data_dict_md(output_dir = tempdir())
```

Add title, detailed descriptions and change default filename

```{R eval=FALSE, echo=TRUE}
description <- data.frame(
                  variable = c("Species"), 
                  description = c("Species of Iris flower"))
data_dict_md(iris, 
             title = "iris flower data set", 
             description =  description, 
             output_file = "data_dict_iris.md",
             output_dir = tempdir())
```

### Basic data cleaning

To clean a variable you can use `clean_var`. With one line of code you can rename a variable, replace NA-values and set a minimum and maximum for the value.

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=3}
iris %>% 
  clean_var(Sepal.Length, 
            min_val = 4.5, 
            max_val = 7.0, 
            na = 5.8, 
            name = "sepal_length") %>% 
  describe()
```

### Connecting to a datawarehouse

The explore package comes with a set easy to remember function to connect, read and write from/to a datawarehouse (dwh) using odbc.

```{R eval=FALSE, echo=TRUE}
# connect to a dwh(odbc DSN must be defined)
dwh <- dwh_connect("DWH_DSN")

# if you need to pass user and password
dwh <- dwh_connect("DWH_DSN", 
                    user = "myuser",
                    pwd = rstudioapi::askForPassword()
                  )

# read table from a dwh
data <- dwh_read_table(dwh, "db.tablename")

# read data from a dwh using sql
data <- dwh_read_data(dwh, sql = "select * from db.tablename")

# disconnect from dwh
dwh_disconnect(dwh)

```

To write large data to a dwh you can use `dwh_fastload()`. It connects to a dwh, writes the data and disconnects.

```{R eval=FALSE, echo=TRUE}
# connect to a dwh(odbc DSN must be defined)
data  %>% dwh_fastload("DWH_DSN", "db.tablename")
```
