---
title: "Predict"
author: "Roland Krasser"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Predict}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The explore package offers a simplified way to use machine learning and make a prediction.

* `explain_tree()` creates a decision tree
* `explain_forest()` creates a random forest
* `explain_logreg()` creates a logistic regression
* `predict_target()` uses a model to make a prediction

We use synthetic data in this example

```{r message=FALSE, warning=FALSE}
library(dplyr)
library(explore)

train <- create_data_buy(obs = 1000, seed = 1)
glimpse(train)
```

### Train model

First we create a decision tree model, using `buy` as target (`buy` contains only 0 and 1 values)

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
train %>% explain_tree(target = buy)
```

We see some clear patterns. Now we create a random forest model (as it is more accurate).
To get the model itself, use parameter `out = "model"`

```{r fig.height=4, fig.width=6, message=FALSE, warning=FALSE}
model <- train %>% explain_forest(target = buy, out = "model")
```

### Predict

Now we create test data and use the model for a prediction. We use a different seed so we get different data.

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
test <- create_data_buy(obs = 1000, seed = 2)
glimpse(test)
```

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
test <- test %>% predict_target(model = model)
glimpse(test)
```

Now we got 2 new variables `prediction_0` (the probability of `buy == 0`) and 
`prediction_1` (the probability of `buy == 1`). We can check the predictions by comparing `prediction_1` with real values of buy.

```{r message=FALSE, warning=FALSE, fig.width=6, fig.height=4}
test %>% explore(prediction_1, target = buy)
```

There is a clear difference between `buy == 0` and `buy == 1`. So the prediction works.
