---
title: "Tips and tricks"
author: "Roland Krasser"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Tips & tricks}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(dplyr)
library(explore)
```

## Count with percent

A classic `count()` returns the number of observations.

```{r echo=TRUE, message=FALSE, warning=FALSE}
data <- use_data_penguins()
data %>% count(island)
```

To add percent values, simply use `count_pct()` from {explore}.

```{r echo=TRUE, message=FALSE, warning=FALSE}
data |> count_pct(island)
```

## Add id

```{r echo=TRUE, message=FALSE, warning=FALSE}
data %>% glimpse()
```

To add an id variable, simply use `add_var_id()` from {explore}.

```{r echo=TRUE, message=FALSE, warning=FALSE}
data %>% add_var_id() %>% glimpse()
```

## User defined report

Create a user defined report (RMarkdown template) to explore your own data.

```{r echo=TRUE, message=FALSE, warning=FALSE}
create_notebook_explore(
  output_dir = tempdir(),
  output_file = "notebook-explore.Rmd")
```

## Basic data cleaning

### Rename variable

```{r echo=TRUE, message=FALSE, warning=FALSE}
data <- use_data_titanic(count = FALSE)
glimpse(data)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
data <- data %>% clean_var(Age, name = "age")
glimpse(data)
```

### Replace NA values

```{r echo=TRUE, message=FALSE, warning=FALSE}
data <- use_data_beer()
data %>% describe(energy_kcal_100ml)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
data <- data %>% clean_var(energy_kcal_100ml, na = 42)
data %>% describe(energy_kcal_100ml)
```

### Set min max values

```{r echo=TRUE, message=FALSE, warning=FALSE}
data <- create_data_person()
data %>% describe(age)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
data <- data %>% clean_var(age, min_val = 20, max_val = 80)
data %>% describe(age)
```

### Rescale 0 to 1

```{r echo=TRUE, message=FALSE, warning=FALSE}
data %>% describe(income)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
data <- data %>% clean_var(income, rescale01 = TRUE)
data %>% describe(income)
```

### Cleaning text

```{r echo=TRUE, message=FALSE, warning=FALSE}
data[1, "handset"] <- " android "
data[2, "handset"] <- "ANDROID"
data %>% describe(handset)
```
```{r echo=TRUE, message=FALSE, warning=FALSE}
data <- data %>% clean_var(handset, simplify_text = TRUE)
data %>% describe(handset)
```
